The Units of Permissionless Consensus: \\ Towards Mobile and Edge Computing

### 2022-10-22 theoretical-consensus/article/article.tex
add page 04

At this point, we can exercise the imagination and extrapolate the
previous block generation mechanism to a \emph{Proof-of-Something} 
pseudo-random competition in which an entity in possession of a higher
amount of a certain resource, either computational power, or stake, 
or certain currency, or a higher amount of storage space, etc., guarantees
a higher probability of leading the block generation and proposal, and
consequently winning acceptance by the majority. This is the essence of
the Proof-of-Stake consensus protocol, as a derivative of the Proof-of-Work
mechanism, where a stake is a traceable and verifiable amount of a certain
unit, token or currency, that is owned by a certain entity that wishes to
participate in the consensus protocol. The stake works as a form of collateral
that is used to guarantee the honesty of the entities, in an attempt to
reduce the Sybil attack probability. And, generally as in Proof-of-Work, 
the higher the stake, the higher the probability of leading the block generation 
and proposal. Proof-of-Stake is, indeed, a more energy efficient consensus protocol,
however, this adds up to nothing, especially when compared to the already proved
BFT solutions, if centralization is a functional requirement, 
or unfair advantage is a potential trait. 
These limitations, along with other security concerns, may push away the 
participation of resourceless entities, and consequently, the network
decentralization.

@@

Due to the self-contained nature of this type of protocols, where the
attestation of the stake is done inside the network, multiple classes
of Proof-of-Stake can be distinguished, by the way the meaningfulness of
the stake is defined and employed. For example, there are the simplest forms of \emph{PoS} that inherit most
of the characteristics of the Proof-of-Work mechanism, but with the
stake as part of the block generation and hashing proof (See Algorithm \ref{alg:PoSBlockGeneration}).
These are commonly referred to as Chain-based \emph{PoS}, inheriting also the
fault-tolerance properties of the Nakamoto consensus protocol.
Committee-based \emph{PoS} protocols, on the other hand, are more complex,
and they are based on the idea of a committee of entities that are 
selected to lead the block generation and proposal. The committee is
formed by a subset of the network entities, and the selection is done
based on their stake. Usually, the selection takes a form of a random
sequence of stake holders, where the probability of being selected
for every spot is proportional to the stake. That sequence is then 
consciously timed to orderly pace the block generation and proposal, 
by the current slot leader. Examples of this type of mechanisms are
the \emph{Ouroboros} \cite{kiayias2017ouroboros}, \emph{Ouroboros Praos} \cite{david2018ouroboros} protocols
or \emph{Snow White} \cite{daian2019snow}. The longest chain rule is
still applied, with probabilistic finality, and the 1/2 threshold is
still required or tolerated for the network to reach consensus. Nonetheless,
Committee-based \emph{PoS} protocols may not pair well with decentralization and 
increase in the committee size, which detereorates the performance of the
protocol. Several improvements have been proposed to tackle not only this
issue, but also the security of the committees which, smaller, could be more
vulnerable to targeted attacks \cite{david2018ouroboros}.

@@

\subsection{}

>>>>

At this point, we can exercise the imagination and extrapolate the
previous block generation mechanism to a \emph{Proof-of-Something} 
pseudo-random competition in which an entity in possession of a higher
amount of a certain resource, either computational power, or stake, 
or certain currency, or a higher amount of storage space, etc., guarantees
a higher probability of leading the block generation and proposal, and
consequently winning the acceptance by the majority. This is the essence of
the Proof-of-Stake consensus protocol, as a derivative of the Proof-of-Work
mechanism, where a stake is a traceable and verifiable amount of a certain
unit, token or currency, that is owned by a certain entity that wishes to
participate in the consensus protocol. The stake works as a form of collateral
that is used to guarantee the honesty of the entities, in an attempt to
reduce the Sybil attack probability. And, generally as in Proof-of-Work, 
the higher the stake, the higher the probability of leading the block generation 
and proposal. Proof-of-Stake is, indeed, a more energy efficient consensus protocol,
however, this adds up to nothing, especially when compared to the already proved
BFT solutions, if centralization is a functional requirement, 
or unfair advantage is a potential trait. 
These limitations, along with other security concerns, may push away the 
participation of resourceless entities, and consequently, affect the network
decentralization and security.

@@

Due to the self-contained nature of this type of protocols, where the
attestation of the stake is done inside the network, multiple classes
of Proof-of-Stake can be distinguished, by the way the meaningfulness of
the stake is defined and employed. For example, there are the simplest forms of \emph{PoS} that inherit most
of the characteristics of the Proof-of-Work mechanism, but with the
stake as part of the block generation and hashing proof (See Algorithm \ref{alg:PoSBlockGeneration}).
These are commonly referred to as Chain-based \emph{PoS}, inheriting also the
fault-tolerance properties of the Nakamoto consensus protocol.
Committee-based \emph{PoS} protocols, on the other hand, are more complex,
and based on the idea of a committee of entities that are 
selected to lead the block generation and proposal. The committee is
formed by a subset of the network entities, and the selection is done
based on their stake. Usually, the selection takes the form of a random
sequence of stake holders, where the probability of being selected
for every spot is proportional to the stake. That sequence is then 
consciously timed to orderly pace the block generation and proposal, 
by the current slot leader. Examples of this type of mechanisms are
the \emph{Ouroboros} \cite{kiayias2017ouroboros}, \emph{Ouroboros Praos} \cite{david2018ouroboros} protocols
or \emph{Snow White} \cite{daian2019snow}. The longest chain rule is
still applied, with probabilistic finality, and the 1/2 threshold is
still required or tolerated for the network to reach consensus. Nonetheless,
Committee-based \emph{PoS} protocols may not pair well with decentralization and 
increase in the committee size, which detereorates the performance of the
protocol. Several improvements have been proposed to tackle not only this
issue, but also the security of the committees which, smaller, could be more
vulnerable to targeted attacks \cite{david2018ouroboros}.

@@

\subsection{Proof-of-Something Else}

Idealized and inspired by Proof-of-Stake, extending or adapting Proof-of-Work
became a popular trend in the blockchain community. The main idea is to
replace the computational power with some other resource, that is
more scarce, or more valuable, or more verifiable, or more traceable, etc.,
or to combine multiple resources, or even to add extra requirements to 
pure Proof-of-Work. From Proof-of-Meaningful-Work, to Proof-of-Work-Time, there
are countless and exotic variations of these mechanisms, and the list is growing \cite{token-economy-gitbook}.

\subsubsection{Proof-of-Space} established itself as a popular alternative among those.
The goal is to replace computational power with storage space,
and if possible give it a meaningful usage. Proof-of-Retrievability is
a cited example of this type of protocols, where the storage space is
used to store a large amount of relevant or public data, and the retrieval of that data
is used as a proof of the storage usage \cite{juels2007pors}. Despite the reciclability 
of the storage space, as a resource that could be a good alternative
to computational power, this algorithm has its own fatal flaws, depending
mostly on a central dealer for randomness as well as for consensus fairness.
Other reported Proof-of-Space variants may not even attempt to store/retrieve any meaningful data.

\subsubsection{Proof-of-Authority} another well known derivative,
which, in effect, is a form of Proof-of-Stake, where the stake is the 
identity of the entity, that has been verified, scrutinized and
approved by the network, giving it the right to participate in the
closed consensus group. The consensus protocol, if not a BFT protocol, 
can follow the style of a Proof-of-Work protocol, depending on the 
requirements of the network - a choice that may shape the fault tolerance,
decentralization, security and performance of the network \cite{survey-dist-consensus}.

\subsubsection{Proof-of-Elapsed-Time} is another alternative to Proof-of-Work,
especially targetting the intensive mining issue, that has been proposed 
by Intel. The idea is to simulate the mining process by using a
randomized time delay, which is proportional to the amount of work
that would be required to mine the block. The time delay is then used to pace the
block generation and proposal, and the block acceptance can be performed
using a BFT protocol \cite{olson2018sawtooth}. This solution, however,
is not a viable alternative to other, permissioned and permissionless, approaches, 
by holding strong and centralized hardware requirements, 
as the computations require a trusted execution environment
(TEE) to establish the trust among the participants. Some efforts have been made
to combine the TEE feature with Proof-of-Stake, in order to achieve a
more decentralized and secure protocol, but the 
same hardware requirements are still a major issue.

\subsection{Other approaches}

\subsubsection{The XRP Ledger Consensus Protocol} powers the XRP Ledger in a
low-latency, high throughput, and  BFT fashion. The network itself operates somewhat
diferently from the other blockchain references, and treats transactions as atomic units of 
the ledger. These transactions are collected by the nodes and relayed to the
nodes' UNL (Unique Node List), which is a list of peers that are trusted by that current
entity. At every epoch, multiple rounds are needed to filter transactions and vote 
on the next state of the ledger. This voting happens at the 
UNLs' level, which are expected to be overlapped by a certain percentage,
in order to achieve the artificial vote threshold for globally accepting the transactions.
That threshold defines the fault tolerance of the protocol, and is set to 80\%, which 
bounds this protocol to a 1/5 BFT allowance \cite{schwartz2014ripple, chase2018analysis}. 
Despite that limitation, when compared to other BFT protocols, this protocol seeks to trade 
fault tolerance for performance, by having lower connectivity and less message exchanges. 
However, it is still expected high synchrony within the UNLs. Given these characteristics, the protocol
is expected to be more suitable for permissioned networks, and, in terms of network size,
is seen as not as scalable and decentralized as the permissionless ones \cite{survey-dist-consensus}.
